---
title: "Iteration in the `tidyverse`"
author: "<br>Alexander Poon<br> [alex.poon@tn.gov](mailto:alex.poon@tn.gov)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
    revealjs::revealjs_presentation:
        theme: night
        transition: none
        css: slides.css
---

``` {r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(width = 65)
```

## Warmup

Q: What's the difference between a vector and a list?

&nbsp;

> * Each element of a vector is a single int or double or character or logical or factor, etc.
> * All elements of a vector have to be of the same *mode*
> * Lists can contain arbitrary objects
> * Lists can be arbitrarily nested

## Preamble

```{r}
library(tidyverse)

(districts <- read_csv("data/districts.csv"))
```

## Iteration

Doing the same(ish) thing over and over (without copy + paste).

## Iteration in `dplyr`

Applying the same function to multiple columns:

```{r, eval = FALSE}
districts %>%
    summarise(
        mean(math, na.rm = TRUE),
        mean(ela, na.rm = TRUE)
    )

districts %>%
    mutate(
        enrollment = enrollment/1000,
        expenditures = expenditures/1000
    )
```

## Iteration in `dplyr`

Applying multiple functions to the same column:

```{r, eval = FALSE}
districts %>%
    summarise(
        mean(grad, na.rm = TRUE),
        min(grad, na.rm = TRUE),
        max(grad, na.rm = TRUE)
    )
```

## Iteration in `dplyr`

`dplyr` has a family of `*_all`, `*_at`, and `*_if` functions to do iteration tasks, where `*` is any of the following:

* `select()`
* `rename()`
* `filter()`
* `arrange()`
* `mutate()`
* `summarise()`
* `transmute()`

## Iteration in `dplyr`

`dplyr` has a family of `*_all`, `*_at`, and `*_if` functions to do iteration tasks.

> * `*_all` applies to all columns in a data frame
> * `*_at` applies to a selection of columns
> * `*_if` applies to a selection of columns meeting a condition

## Arguments

`*_all`, `*_if`, and `*_at`, functions take the following arguments:

&nbsp;

|  function  | .tbl | .vars | .predicate | .funs |
|------------|------|-------|------------|-------|
|   *_all    |  x   |       |            |   x   |
|   *_if     |  x   |       |     x      |   x   |
|   *_at     |  x   |   x   |            |   x   |

&nbsp;

* `.tbl`: a data frame
* `.vars`: column(s) in `.tbl`
* `.predicate`: a function to select columns based on values
* `.funs`: one or more functions

. . .

There is tons of flexibility in specifying `.vars` and `.funs`.

## `.vars`

`.vars` can be a character vector:

```{r, eval = FALSE}
districts %>%
    summarise_at(
        .vars = c("math", "ela"),
        .funs = mean
    )
```

## `.vars`

`.vars` can be unquoted variable names, wrapped in `vars()`:

```{r, eval = FALSE}
districts %>%
    summarise_at(
        .vars = vars(math, ela),
        .funs = mean
    )
```

## `.vars`

`vars()` follows the same syntax as `select()`:

```{r, eval = FALSE}
districts %>%
    summarise_at(
        .vars = vars(-system, -system_name, -region),
        .funs = mean
    )
```

Applies `mean` to all columns except `system`, `system_name`, and `region`.

## `.vars`

`vars()` follows the same syntax as `select()`:

```{r, eval = FALSE}
districts %>%
    summarise_at(
        .vars = vars(alg_1:dropout),
        .funs = mean
    )
```

Applies `mean` to all columns positionally from `AlgI` to `Dropout`.

## `.vars`

`vars()` follows the same syntax as `select()`:

```{r, eval = FALSE}
districts %>%
    summarise_at(
        .vars = vars(starts_with("alg")),
        .funs = mean
    )
```

Use select helpers in `vars()`:

* `starts_with()`
* `ends_with()`
* `contains()`
* `matches()` (a regular expression)
* etc.

## `.predicate`

`.predicate` takes a function that returns `TRUE` or `FALSE` based on column values:

```{r, eval = FALSE}
districts %>%
    select_if(is.character)
```

## `.funs`

`.funs` can be a named function:

```{r, eval = FALSE}
districts %>%
    summarise_at(c("math", "ela"), mean)
```

## `.funs`

Pass other arguments to `funs` via `...`:

```{r, eval = FALSE}
districts %>%
    summarise_at(c("math", "ela"), mean, na.rm = TRUE)
```

## `.funs`

`.funs` can be an anonymous function:

```{r, eval = FALSE}
districts %>%
    mutate_at(c("enrollment", "expenditures"), function(x) x/1000)
```

## `.funs`

`.funs` can be an anonymous function:

```{r, eval = FALSE}
districts %>%
    mutate_at(c("enrollment", "expenditures"), ~ ./1000)
```

`~` is shorthand for an anonymous function with one argument.

## `.funs`

`.funs` can be multiple functions, wrapped in `funs`:

```{r}
districts %>%
    summarise_at(
        .vars = c("grad", "dropout"),
        .funs = funs(mean, min, max),
        na.rm = TRUE  # passed to mean, min, max via ...
    )
```

Any `...` arguments will apply to each of `funs`.

## Examples

Count the number of missing values in each column:

. . .

```{r}
districts %>%
    summarise_all(~ sum(is.na(.)))
```

## Examples

Count the number of missing values in each column:

```{r}
districts %>%
    mutate_all(is.na) %>%
    summarise_all(sum)
```

## Examples

Keep all the columns with no missing values:

. . .

```{r}
districts %>%
    select_if(~ all(!is.na(.)))
```

## Examples

Round all numeric columns to the whole number:

. . .

```{r}
districts %>%
    mutate_if(is.numeric, round)
```

## Examples

Take the average of all district profile columns:

. . .

```{r}
districts %>%
    summarise_at(vars(enrollment:expenditures), mean, na.rm = TRUE)
```

## Iteration Pre-`tidyverse`

Vectorized functions:

```{r}
x <- 1:10

x^2
```

## Iteration Pre-`tidyverse`

For loops:

```{r, eval = FALSE}
for (x in vec) {
    # Do something
}
```

. . .

Not very idiomatic in R.

## Iteration Pre-tidyverse

`apply` family in base R:

```{r, eval = FALSE}
lapply(X, fun, ...)
```

Applies a function to each element of a vector or a list `X`.

. . .

Superseded by the `map` family in `purrr`.

## The `purrr::map` family

A family of functions that take a list (or vector), applies a function to each element, and provides some flexibility in how the result is returned.

<img src="images/purrr.svg" width="200">

## Scenario

I build models to predict whether students in middle school will graduate high school with a 21+ ACT Composite based on:

* test scores
* absenteeism
* disciplinary issues
* mobility
* the school they attend

## Scenario

```{r}
(grade_8 <- read_csv("data/grade_8.csv"))
```

## Scenario

I train a few candidate models on a training set of one cohort of students, predict on a test set of a second cohort, and check classification accuracy.

| Student | Prediction |  Outcome  |
|---------|------------|-----------|
|    1    |  Ready     | Ready     |
|    2    |  Not Ready | Not Ready |
|    3    |  Not Ready | Not Ready |
|    4    |  Ready     | Not Ready |
|    .    |     .      |    .      |
|    .    |     .      |    .      |
|    .    |     .      |    .      |

## Preamble 2

Read in models and test set:

```{r}
library(caret)

# Pre-trained models
glm <- read_rds("models/glm_8.rds")
nnet <- read_rds("models/nnet_8.rds")
rpart <- read_rds("models/rpart_8.rds")
xgb <- read_rds("models/xgb_8.rds")

# Test set to make predictions
test_set <- filter(grade_8, cohort == 2012)
```

## Preamble 2

Split and preprocess test set:

```{r}
predictors <- c("absences", "enrollments", "expulsions", "suspensions",
    "math", "read", "school_math", "school_read")

test_x <- test_set[predictors]
test_y <- factor(test_set$ready_grad)

# Models were trained on centered/scaled data
# Center and scale test predictors
test_preprocess <- preProcess(test_x, method = c("center", "scale"))

test_x <- predict(test_preprocess, test_x)
```

## Example

I want to do this:

```{r, eval = FALSE}
predict(glm, test_x)
predict(nnet, test_x)
predict(rpart, test_x)
predict(xgb, test_x)
```

. . .

where:

* `glm` is logistic regression
* `nnet` is a simple neural network
* `rpart` is a decision tree
* `xgb` is (eXtreme) gradient boosting

## Example

With `map`:

```{r, eval = FALSE}
map(.x = list(glm, nnet, rpart, xgb), .f = predict, test_x)
```

Iterates by applying function `.f` over each element of `.x`, with the common argument `test_set`.

## Example

With `map`:

```{r}
map(.x = list(glm, nnet, rpart, xgb), .f = predict, test_x) %>% 
    str()
```

This returns a list, where each element is the result of predicting on the test set with the appropriate model.

## `map`

If `.x` is a named list, those names will persist through `map`s:

```{r}
model_list <- list(
    "glm" = glm, 
    "nnet" = nnet, 
    "rpart" = rpart, 
    "xgb" = xgb
)

map(.x = model_list, .f = predict, test_x) %>%
    str()
```

## `map`

`map`s are composable with `%>%`:

```{r, eval = FALSE}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map(mean)
```

## `map`

`map`s are composable with `%>%`:

```{r}
model_list %>%
    map(predict, test_x) %>%
    str()
```

The first `map` takes each model and predicts on the test set, returning a 10,000-length vector of `ready`/`not ready` for each model.

## `map`

`map`s are composable with `%>%`:

```{r}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    str()
```

The second `map` takes each list of predictions and compares it element-wise to `test_y`, returning a 10,000-length vector of `TRUE`/`FALSE` for each model.

## `map`

`map`s are composable with `%>%`:

```{r}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map(mean)
```

The third `map` averages each list of booleans, giving a proportion of correct predictions.

## `map_*`

If the result of each iteration is a length 1 vector of the same type, `map_*` will return an atomic vector of type `*`:

```{r, eval = FALSE}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map_dbl(mean)
```

where `*` can be any of `lgl`, `chr`, `int`, `dbl`.

## `map_*`

If the result of each iteration is a length 1 vector of the same type, `map_*` will return an atomic vector of type `*`:

```{r}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map_dbl(mean)
```

. . .

This is useful because vectors are easier to work with than lists, and it's easier to reason about code when you know what it returns.

## `map_*`

`map_*` returns an error if the result is not of the expected type:

```{r, error = TRUE}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map_lgl(mean)
```

## `map_*`

There are also `map_dfc` and `map_dfr` variants, which bind the list elements as columns/rows of a data frame respectively:

```{r}
model_list %>%
    map_dfc(predict, test_x)
```

## `.f` in `map`

`.f` follows all the conventions of the `dplyr` iteration functions:

```{r, eval = FALSE}
model_list %>%
    map(predict, test_x) %>%
    map(~ . == test_y) %>%
    map_dbl(mean)
```

* Can be a named function
* Can be an anonymous function
* Can use the `~` shorthand

## `.f` in `map`

`.f` follows all the conventions of the `dplyr` iteration functions:

```{r, eval = FALSE}
model_list %>%
    map(predict, test_x, type = "prob")
```

* Take additional arguments with `...`

## `.f` in `map`

`.f` follows all the conventions of the `dplyr` iteration functions:

```{r}
model_list %>%
    map(predict, test_x, type = "prob") %>% 
    str()
```

## `.f` in `map`

`.f` can be an integer or a character string, in which case it will extract an element by position (integer) or name (string).

```{r}
model_list %>%
    map(predict, test_x, type = "prob") %>%
    map("ready") %>%
    str()
```

## `.f` in `map`

Calculate area under the [Receiver Operating Characteristic curve](https://en.wikipedia.org/wiki/Receiver_operating_characteristic):

```{r}
model_list %>%
    map(predict, test_x, type = "prob") %>%
    map("ready") %>%
    map_dbl(~ yardstick::roc_auc_vec(truth = test_y, estimate = .))
```

## Preamble 3

I've decided to use `nnet`, and trained a model on each of three cohorts, and want to predict on each of three test sets.

. . .

I want to do something like this:

```{r, eval = FALSE}
predict(nnet_6, test_x_6)
predict(nnet_7, test_x_7)
predict(nnet_8, test_x_8)
```

. . .

i.e., iterating over two arguments instead of one.

## Preamble 3

Read in trained model and test set for each grade:

```{r}
nnet_list <- list(
    "grade_6" = read_rds("models/nnet_6.rds"),
    "grade_7" = read_rds("models/nnet_7.rds"),
    "grade_8" = read_rds("models/nnet_8.rds")
)

test_list <- list(
    "test_6" = read_csv("data/grade_6.csv") %>% filter(cohort == 2012),
    "test_7" = read_csv("data/grade_7.csv") %>% filter(cohort == 2012),
    "test_8" = read_csv("data/grade_8.csv") %>% filter(cohort == 2012)    
)
```

## Preamble 3

Write a function to preprocess test set predictors:

```{r}
preprocess_test <- function(cohort) {

    predictors <- c("absences", "enrollments", "expulsions", "suspensions",
        "math", "read", "school_math", "school_read")

    x <- cohort[predictors]

    train_preprocess <- preProcess(x, method = c("center", "scale"))

    predict(train_preprocess, x)

}
```

## Preamble 3

Split test sets into (preprocessed) predictors and outcomes:

``` {r}
test_x <- map(.x = test_list, .f = preprocess_test)
test_y <- map(.x = test_list, .f = "ready_grad")
```

## `map2`

The solution is `map2`:

```{r, eval = FALSE}
map2(.x = nnet_list, .y = test_x, .f = predict)
```

`.x` and `.y` are lists of the same length, which will be cycled through as the first and second arguments of `.f`.

## `map2`

The solution is `map2`:

```{r}
map2(.x = nnet_list, .y = test_x, .f = predict) %>% 
    str()
```

## `map2`

`map2` behaves otherwise identically to `map`:

```{r, eval = FALSE}
map2(.x = nnet_list, .y = test_x, .f = predict) %>%
    map2(.x = ., .y = test_y, .f = ~ .x == .y) %>%
    map_dbl(mean)
```

> * Composable with pipes
> * Take named or anonymous functions
    * For a two-argument anonymous function, use `.x` and `.y`
> * Take additonal `.f` arguments with `...`
> * Has `_int`, `_chr`, `_dbl`, `_lgl` variants

## `map2`

`map2` behaves otherwise identically to `map`:

```{r}
map2(.x = nnet_list, .y = test_x, .f = predict) %>%
    map2(.x = ., .y = test_y, .f = ~ .x == .y) %>%
    map_dbl(mean)
```

## `map` with Arbitrary Number of Arguments

For iterating over arbitrary number of arguments, use `pmap`:

```{r, eval = FALSE}
pmap(
    .l = list(
        c("A", "B", "C"),
        c(1, 2, 3),
        c(foo, bar, baz)
    ),
    .f = function(x, y, z) {
        # do something
    }
)
```

`pmap` takes a list of lists (or vectors), and calls `.f` with the first element of each sublist as its arguments, then the second element of each sublist, etc.

## `map` with Arbitrary Number of Arguments

For iterating over arbitrary number of arguments, use `pmap`:

```{r, eval = FALSE}
pmap(
    .l = list(
        models = nnet_list
        test_sets = test_x
    ),
    .f = predict
)
```

## `map` with data frames

Recall:

A data frame is a list of vectors.  
A tibble is a list of lists.

&nbsp;

. . .

Use `map_*` to apply functions over columns of a data frame, e.g.:

```{r}
map_chr(test_set, class)
```

## `map` on data frames

You can use `map`s on list-columns of a `tibble`:

```{r}
(model_df <- tibble(
    grade = c(6, 7, 8),
    test_set = test_x,
    model = nnet_list
))
```

## `map` on data frames

You can use `map`s on list-columns of a `tibble`:

```{r}
map2(.x = model_df$model, .y = model_df$test_set, .f = predict) %>%
    str()
```

## `map` on data frames

You can use `map` in conjunction with `mutate`:

```{r, eval = FALSE}
model_df %>%
    mutate(
        predictions = map2(.x = model, .y = test_set, .f = predict),
        accuracy = map2(.x = model, .y = test_set, .f = predict) %>%
            map2(.x = ., .y = test_y, .f = ~ .x == .y) %>%
            map_dbl(mean)
    )
```

. . .

```{r, echo = FALSE}
model_df %>%
    mutate(
        predictions = map2(.x = model, .y = test_set, .f = predict),
        accuracy = map2(.x = model, .y = test_set, .f = predict) %>%
            map2(.x = ., .y = test_y, .f = ~ .x == .y) %>%
            map_dbl(mean)
    )
```

## `map` on data frames

You can replicate the iteration functions in `dplyr` with `map`:

```{r}
summarise_all(districts, ~ sum(is.na(.)))
```

## `map` on data frames

You can replicate the iteration functions in `dplyr` with `map`:

```{r}
map_int(districts, ~ sum(is.na(.)))
```

## `map`

You can `map` all manner of repetitive code:

```{r, eval = FALSE}
glm <- read_rds("models/glm_8.rds")
nnet <- read_rds("models/nnet_8.rds")
rpart <- read_rds("models/rpart_8.rds")
xgb <- read_rds("models/xgb_8.rds")
```

. . .

becomes:

```{r, eval = FALSE}
model_list <- map(
    .x = c("models/glm_8.rds", "models/nnet_8.rds", "models/rpart_8.rds", "models/xgb_8.rds"),
    .f = read_rds
)
```

## List Unpacking

If you want to assign the components of a `map` to different objects, use the multiple assignment operator (`%<-%`) from `zeallot`:

```{r, eval = FALSE}
c(glm, nnet, rpart, xgb) %<-% map(
    .x = c("models/glm_8.rds", "models/nnet_8.rds", "models/rpart_8.rds", "models/xgb_8.rds"),
    .f = read_rds
)
```

## List Unpacking

In general, you should keep things as a list unless you're going to do the same things with each element.

```{r}
model_list <- map(
    .x = c("models/glm_8.rds", "models/nnet_8.rds", "models/rpart_8.rds", "models/xgb_8.rds"),
    .f = read_rds
)
```

. . .

vs.

```{r, eval = FALSE}
c(train_8, test_8) %<-% split(grade_8, grade_8$cohort)
```

## `map`

Iterate over large blocks of code succinctly with `map`:

e.g.: for each grade:

* read in the data
* split into training and test sets
* split training and test sets into features and outcome
* center and scale features
* fit a model

## `map`

To train models, I do something like this for each grade:

```{r, eval = FALSE}
train_set <- read_csv("data/grade_8.csv") %>%
    filter(cohort == 2011)

predictors <- c("absences", "enrollments", "expulsions", "suspensions",
    "math", "read", "school_math", "school_read")

train_x <- train_set[predictors]
train_y <- train_set$ready_grad

center_scale <- preProcess(train_x, method = c("center", "scale"))

train_x <- predict(center_scale, train_x)

train(x = train_x, y = train_y, method = "nnet", metric = "ROC")
```

## `map`

Turn it into a function:

```{r, eval = FALSE}
fit_nnet <- function(grade) {
    
    train_set <- read_csv(grade) %>%
        filter(cohort == 2011)

    predictors <- c("absences", "enrollments", "expulsions", "suspensions",
        "math", "read", "school_math", "school_read")

    x <- grade[predictors]
    y <- grade$ready_grad

    center_scale <- preProcess(x, method = c("center", "scale"))

    x <- predict(center_scale, x)

    train(x = x, y = y, method = "nnet", metric = "ROC")

}
```

## `map`

Iterate with `map`:

```{r, eval = FALSE}
map(
    .x = c("data/grade_6.csv", "data/grade_7.csv", "data/grade_8.csv"),
    .f = fit_nnet
)
```

## `walk`

In programming, *side effects* refer to anything a function does other than providing a return value.

> * `write_csv()`
> * `print()`
> * `plot()`
> * `library()`

## `walk`

`walk` is the equivalent of `map` for producing side effects.

```{r, eval = FALSE}
walk2(
    .x = model_list,
    .y = paste0("models/", names(model_list), "_8.rds"),
    .f = write_rds
)
```

## `reduce`

From the docs:

> Reduce a list to a single value by iteratively applying a binary function:

. . .

`ach`, `growth`, etc. are data frames to be joined:

``` {r, eval = FALSE}
ach %>%
    left_join(growth, by = c("system", "school")) %>%
    left_join(grad, by = c("system", "school")) %>%
    left_join(ready_grad, by = c("system", "school")) %>%
    left_join(absenteeism, by = c("system", "school")) %>%
    left_join(elpa, by = c("system", "school"))
```

## `reduce`

With `reduce`, this becomes:

```{r, eval = FALSE}
reduce(
    .x = list(ach, growth, grad, ready_grad, absenteeism, elpa),
    .f = left_join,
    by = c("system", "school")  # passed to left_join via ...
)
```

## Further Reading

* A `purrr` DataCamp tutorial: https://www.datacamp.com/courses/writing-functions-in-r
* Charlotte Wickham's `purrr` tutorial for working with nested lists: https://github.com/cwickham/purrr-tutorial
* The `purrr` cheat sheet: https://www.rstudio.com/resources/cheatsheets/#purrr
