---
output:
    word_document:
        reference_docx: document_template.docx
---

``` {r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(readr)
library(tidyr)
library(dplyr)

ach_profile <- read_csv("data/achievement_profile_data_with_CORE.csv")
```

## Davidson County District Performance

**Introduction**

&nbsp;

Davidson County is a school district serving `r ach_profile[ach_profile$system_name == "Davidson County", ]$Enrollment` students. Its expenditures in 2015 were $`r sprintf("%.2f", ach_profile[ach_profile$system_name == "Davidson County", ]$Per_Pupil_Expenditures)` per student.

&nbsp;

**Demographic Information**

&nbsp;

Here is a table with student demographic information for Davidson County:

&nbsp;

``` {r}

ach_profile %>%
    filter(system_name %in% c("Davidson County", "State of Tennessee")) %>%
    select(system_name, Pct_BHN, Pct_ED, Pct_SWD, Pct_EL) %>%
    rename(`Black/Hispanic/Native American` = Pct_BHN, `Economically Disadvantaged` = Pct_ED, `Students with Disabilities` = Pct_SWD, `English Learners` = Pct_EL) %>%
    gather(Subgroup, `Percentage of Students`, 2:5) %>%
    spread(system_name, `Percentage of Students`) %>%
    mutate(`Davidson County` = paste0(`Davidson County`, "%"),
        `State of Tennessee` = paste0(`State of Tennessee`, "%")) %>%
    knitr::kable()

```
